### 2018-08-24 {{ cookiecutter.repo_name }}/Section_files/06_acknowledgements.tex

%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Main MNRAS template for Cookicutter project
%%
%% Section -- ACKWNOLEDGEMENTS
%%
%% Author: Victor F. Calderon (http://vcalderon.me)
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% In here, you will add the 'abstract' of your paper

\section{Acknowledgements}
\label{sec:acknowledgements}

This is where the acknowledgements go.
It is useful to use \textbf{Acknowledgement Generator} (\https{https://astrofrog.github.io/acknowledgment-generator/}) to create your
acknowledgements. This lets you add the bibitem entried for each of the
packages used in the analysis.